---
title: "plan"
author: "Dan Kelley"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 5
    fig_height: 5
    dpi: 72
    dev.args: list(pointsize=11)

vignette: >
  %\VignetteIndexEntry{oce}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- edit vignettes/plan.Rmd then run with -->
<!-- devtools::build_vignettes() -->
<!-- in the console. -->

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

**Abstract.** The `plan` package provides functions for planning and
describing projects and monitoring the progress towards completions of
individual tasks within projects.

# Introduction

This package provides (mainly graphical) tools for the planning of projects.
This is an early version of the package, providing support for burn-down
charts and Gantt diagrams.


# Burn-down charts

Burndown
charts^[http://alistair.cockburn.us/crystal/articles/evabc/earnedvalueandburncharts.htm]
are used to display a time series of the progress towards the goals of a
project.  The format is simple.  The x axis represents time, ranging from the
onset of the project to the deadline for completion.  The y axis represents the
remaining effort that is require to accomplish the work. As the work is carried
out, the chart provides a running summary of progress towards the deadline.  If
work is proceeding smoothly according to schedule, the chart takes the form of
a triangle, with the remaining effort falling from its initial value to zero at
the deadline. For guidance, a gray line is drawn to indicate this ideal
situation. If work is being accomplished faster than expected, the data will
lie below this gray line. However, if the project is falling behind schedule,
the data line will lie above the ideal line.  Thus, a glance at the chart
indicates whether the deadline can be met.  In most cases, the work is divided
into sub-tasks, and the remaining effort in each task is shown with a different
colour in the chart. This is helpful in identifying tasks that may need more
attention.

The following plots the burndown chart
```{r}
library("plan")
data(burndown)
plot(burndown)
```
To work with your own files, use `read.burndown()`.

# Gantt charts

Gantt diagrams^[http://en.wikipedia.org/wiki/Gantt_Chart] indicate a timetable
for completion of the components of a project. The sample dataset `gantt` is a
hypothetical research plan for an MSc project; you can see how it is graphed
with `demo(gantt)` or `example(plot.gantt)`.

The following plots the built-in Gantt diagram. If this seems promising, the
reader should continue reading, to learn how to make objects of `gantt-class`.
If not, the best is to try some other software, because only minimal
customization is permitted by the `plot,gantt-method`.

```{r}
library("plan")
data(gantt)
plot(gantt)
```

There are three ways to create objects of `gantt-class`: (a) write information
in a text file and use `read.gantt()`, (b) use `as.gantt()` to assemble the
object in one (somewhat complicated) step, or (c) use `new()` to create an
object and then `ganttAddTask()` to add tasks one by one. The third approach
may be the easiest for beginners, so it is illustrated below, with a sketch of
a typical oceanography MSc program.


```{r}
library("plan")
g <- new("gantt")
g <- ganttAddTask(g, "Courses") # no times, so a heading
g <- ganttAddTask(g, "Physical Oceanography", "2016-09-03", "2016-12-05")
g <- ganttAddTask(g, "Chemistry Oceanography", "2016-09-03", "2016-12-05")
g <- ganttAddTask(g, "Fluid Dynamics", "2016-09-03", "2016-12-05")
g <- ganttAddTask(g, "Biological Oceanography", "2017-01-03", "2017-04-05")
g <- ganttAddTask(g, "Geological Oceanography", "2017-01-03", "2017-04-05")
g <- ganttAddTask(g, "Geophysical Fluid Dynamics", "2017-01-03", "2017-04-05")
g <- ganttAddTask(g, "Research") # no times, so a heading
g <- ganttAddTask(g, "Literature review", "2016-09-03", "2017-04-05")
g <- ganttAddTask(g, "Develop analysis skills", "2016-09-03", "2017-08-01")
g <- ganttAddTask(g, "Thesis work", "2017-01-01", "2018-04-01")
g <- ganttAddTask(g, "Defend thesis proposal", "2017-05-01", "2017-06-01")
g <- ganttAddTask(g, "Write papers & thesis", "2017-05-01", "2018-04-01")
g <- ganttAddTask(g, "Defend thesis", "2018-05-01", "2018-05-15")
## Set 'font' to highlight headings
font <- rep(1, length(g[["key"]]))
font[c(1,8)] <- 2
plot(g, ylabel=list(font=font))
```

